### 2018-04-14 ident/docs/draft_2.tex
complete results except noise section

	\subsection{Identifying parameters in kinetic models of metabolism: an example}\label{sec:example}	
	In this section, we illustrate the proposed methodology to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network} and Section \ref{sec:small-model} of the Supplementary Information). We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) for this demonstration. 
	\begin{align}\tag{S5}
	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
	\end{align}		
	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ for $v_1$, as mentioned in Section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments. So, from the $n_E = 21$ different experiments described in Section \ref{sec:experiments} and Table \ref{tab:pval}, we can choose multiple combinations of $p = 2$ experiments to satisfy the data requirements for identifying $v_1$ i.e., in Equation (\ref{eq:nonlineq}) $j = \{1, 2\}$. We label the available concentrations and fluxes as ${ac}^{(j)}$ and ${v_1}^{(j)}$, respectively. Then, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for $v_1$ as:
	\begin{align*}%\label{eq:nonlin-flux2}
	{v_1}^{(j)} = V_{1}^{max}\frac{ac^{(j)}}{ac^{(j)}+K_{1}^{ac}(ne)} &&  j=\{1, 2\}
	\end{align*}

	Solving this simultaneous system of equations in $\mathbb{R}^2$ using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b). These expressions have the form shown in Equation (\ref{eq:theta-eq}).
	\begin{subequations}\label{eq:v1_par}
		\begin{align}		
		\theta_1 &= V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
		\theta_2 &= K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
		\end{align}
	\end{subequations}
	In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b) for different available experimental data combinations. Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b). 
	We can also apply the proposed methodology to practically identify parameters in $v_1$ under the assumption that the protein concentration for the enzyme \textit{E} is available, in addition to the measured concentration of the metabolites and fluxes. In doing so, we get two expressions similar to the one shown in Equation (\ref{eq:v1_par}) for $k_1^{cat}$ and $K_1^{ac}$. The globally identifiable parameter values for all fluxes modeled as Michaelis-Menten equations in provided in Figure \ref{fig:parameter_value_proof}a and \ref{fig:parameter_value_proof}b.

	We apply the method described in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, and illustrated above for $v_1$, to other fluxes of the network (Figure \ref{fig:network}) not described using Michaelis-Menten equations. On applying this method to $v_3$, which is modeled using MWC kinetics, we find that we could not apply our methodology to determine closed form solution for all the parameters in $v_3$. However, when $v_3$ is instead modeled using Convenience kinetics, we find that the parameters are locally identifiable based on the definition we provide in Section \ref{sec:ident}. All three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ used to model $v_3$ have two possible roots. We still could not determine identifiable values for these parameters (Figure \ref{fig:parameter_value_v3}) using the constraints on the physical relevance of their values (see Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}). In the box plot in Figure \ref{fig:parameter_value_v3}a and b we show all the possible estimated parameter values for both roots of all three parameters modeling $v_3$, when experimental data is generated using the MWC model for $v_3$. Due to the numerous estimated parameter values, we can designate these parameters are practically non-identifiable. Although using the Convenience kinetics model for $v_3$ to generate the in silico experimental data leads to a reduction in the uncertainty of the estimated parameters (\ref{fig:parameter_value_v3}c and d), we still end up with non-unique parameter estimates. This leads is to conclude that under the current conditions, flux $v_3$ is structurally locally identifiable and practically non-identifiable.

	However, $K_3^{fdp}$ and $V_3^{max}$ become globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value, and $K_3^{pep}$ and $V_3^{max}$ become globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (Figure \ref{fig:parameter_value_v3_var}).

	Using Figure \ref{fig:parameter_value_v3_var}, we also want to bring attention to the fact that parameter non-identifiability remains when there is a mismatch between the kinetic rate law used to generate the in silico data, and the kinetic rate law used to identify parameters. For example, in the specific case of $v_3$, we see that globally structurally identifiable parameters $V_3^{max}$ and $K_3^{fdp}$ (when $K_3^{pep}$ is fixed) are practically non-identifiable when data generated from experiments using the MWC kinetics is used to model $v_3$ (Figure \ref{fig:parameter_value_v3_var}a). The same is true of globally structurally identifiable parameters $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is fixed (Figure \ref{fig:parameter_value_v3_var}b). Instead, when Convenience kinetics is used to generate in silico data, under the aforementioned conditions all the parameters are globally structurally and practically identifiable (Figure \ref{fig:parameter_value_v3_var}c and d).

	Similarly, we applied the demonstrated methodology to flux $v_5$ that describes a transcription/translation reaction using Hill kinetics, based on only the available metabolomics and the fluxomics information. When we tried identifying the three parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$, we could not obtain closed form solutions for these parameters using the computer algebra system. However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$. Thus, similar to $v_3$ earlier, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in section \ref{sec:ident}. In this instance, the identification required data from two different experiments. Again, just as we see for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally identifiable. On the other hand, $V_e^{max}$ has only one possible root, thereby it is a globally structurally identifiable parameter. However, only one of the two roots ($K_e^{fdp}(2)$) is practically identifiable based on the physical relevance criteria stated in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b. Thus, enforcing this criteria results in global practical identifiability of both $V_e{max}$ and $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:parameter_value_proof}c).

	So, in summary, we find that the identifiability of parameters in kinetic models of metabolism using steady state information is dependent on the kinetic rate law used to model the fluxes within metabolism. Although in some instances (e.g., $v_5$) local identifiability could be resolved to obtain global identifiability using constraints on the values of the parameters such that they are physically relevant, this may not be possible in other scenarios (e.g., $v_3$). Moreover, we also see that the structural identifiability of the parameters in any given kinetic rate law model has a bearing on the ability to determine the practical identifiability of parameters using steady state metabolomic, fluxomic and proteomic information. We find that these can sometimes be resolved by reducing the dimension of the parameter space that is being identified: $\theta \in \mathbb{R}^3$ to $\theta \in \mathbb{R}^2$ for both $v_3$ and $v_2$. Additionally, we would also like to point out that discrepancies between in vivo kinetic rate law from which typical experimental data is obtained, and the in vitro rate law used in kinetic models can itself lead to parameter non-identifiability or local identifiability. This can lead to uncertainty in parameter estimates made from in vivo experimental data.

@@

	\subsection{Degree of identifiability of different enzyme kinetic rate laws}\label{sec:ident_rate_laws}		
	In Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}, we show the number and percentage of experimental data combinations that are capable of identifying each parameter in each of the four fluxes $v_1$ (Figure \ref{fig:ident}a and e), $v_2$ (Figure \ref{fig:ident}b and f), $v_3$ (Figure \ref{fig:ident}c and g) and $v_5$ (Figure \ref{fig:ident}d and h) belonging to the network shown in Figure \ref{fig:network}. Based on the definition given in Section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter. We see from Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2} that the degree of identifiability of the maximum reaction rates ($V_i^{max}$), whenever they are identifiable for each of the four fluxes, is greater than or equal to the degree of identifiability of the corresponding enzyme binding ($K_i$) constants, or the allosteric activation constant ($K_3^{pep}$), in the respective reaction rate law models. We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience kinetic rate law and $v_5$ is modeled as a Hill equation with inhibition. This observation is also true irrespective of the kinetic rate law used to model $v_3$ (MWC in Figure \ref{fig:ident}a-d vs Convenience kinetics in Figure \ref{fig:ident}e-h) to generate in silico experimental data. We accordingly surmise that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.

@@

	The difference in the degrees of identifiability between $v_1$ (Figure \ref{fig:ident}a) and $v_2$ (Figure \ref{fig:ident}b) can be attributed to the ability of data from different combinations of experiments to satisfy the conditions determined for practical identifiability of that parameter. To elaborate, we showed in Equation (\ref{eq:v1_par}) and Section \ref{sec:example} that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them. So, in this instance, recall that the ability of a combination of experiments to be useful is determined by the ability of the input change (changes in the input acetate concentration) to effect a change in the measured value of $v_1$. Similarly, the identification of parameters for $v_2$ (Figure \ref{fig:ident}b) requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}. We elaborate on the type and nature of experiments that are useful and informative for parameter identification in $v_1$ and $v_2$ later in Section x.

	The difference in the degree of identifiability of $v_1$ that comes about from the use of experimental data from different rate law models used to model $v_3$ (Figures \ref{fig:ident}a and e) can be again attributed to the usefulness of experimental data. However, the difference in data usefulness between similar experiments arising due to differences in the kinetic rate law used for $v_3$ could be due to the difference in the dynamics of the network represented by either of the two models. The interaction between different components of the network (Figure \ref{fig:network}) as exemplified by the different models plays an important role in determining the dynamic and the steady state response of the network. The difference in dynamics in turn contributes to different experimentally observed values of metabolite concentrations and fluxes (Figure \ref{fig:experimental_data_dist}), which in turn changes the usefulness of the given combination of datasets, and its subsequent effect of parameter identifiability.

	Nonetheless, we would like to note that these variations in the dynamics contributed by the difference between the rate laws can be resolved with more information i.e., experimental proteomic data that provide protein concentrations for different experiments. We earlier saw that relaxing the assumption on the available experiment data, i.e., assuming the availability of protein concentration (\textit{E}), in addition to the metabolite concentrations (\textit{pep} and \textit{fdp}) and fluxes, results in removal of uncertainty and global identifiability of $v_1$ (Supplementary Figure \ref{fig:parameter_value_proof}). Recall that any predicted parameter value for both $k_1^{cat}$ and $V_1^{max}$ matches with the exact known value irrespective of the experiment (MWC vs Convenience kinetics) from which data is obtained when enzyme concentrations are included for identifiability (Supplementary Figure \ref{fig:parameter_value_proof}a).  Additionally, we also see an improvement in the degree of identifiability of the corresponding parameters $k_1^{cat}$ and $K_1^{ac}$ (Supplementary Figure \ref{fig:ident_root2}a and b) when enzyme concentrations are available.

	We would also like to point out that a difference in either the ability to estimate true parameter values, or the the identifiability of parameter modeling $v_2$ does not arise due to difference in the kinetic models used to obtain in silico experimental data. This could because of the fact that any and all available information that is required for the identification of parameters in $v_2$ is encompassed in the steady state data used for identifiability analysis.

	However, unlike $v_2$, we see that the degree of identifiability of both locally identifiable roots of parameters in $v_3$ changes with the change in the data (MWC vs Convenience kinetics) used for the analysis (Figure \ref{fig:ident}c and g). Despite the difference in the dynamics represented by the different data sets, and its impact on the usefulness of data sets for identifiability purposes that was explained earlier, we could also allude to the structural non-identifiability between the parameters in $v_3$ for the impact that it would have on the practical identifiability of the parameters using in silico data from different sources. In order to ascertain this, we plot the degrees of identifiability of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ under conditions when they are globally identifiable (Section \ref{sec:example}: $V_3^{max}$ and $K_3^{fdp}$ are globally identifiable when $K_3^{pep}$ is constant, and $V_3^{max}$, $K_3^{pep}$are globally identifiable when $K_3^{fdp}$ is known and fixed) in Supplementary Figure \ref{fig:ident_v3_var}. Now that structurally non-identifiability between parameters has been alleviated, we see that the uncertainty introduced by the data obtained from the MWC model results in only a fraction of data sets being useful to estimate parameters (Supplementary Figure \ref{fig:ident_v3_var}a and b). This ties back with our earlier observation that these parameters are not practically identifiable from data sets obtained using the MWC model (Supplementary Figure \ref{fig:parameter_value_v3_var}a and b). However, we see from Supplementary Figures \ref{fig:ident_v3_var}c and d that the globally structurally identifiable parameters are also practically identifiable based on the data obtained from the Convenience kinetics model. This can be corroborated by the parameter values shown in Supplementary Figure \ref{fig:parameter_value_v3_var}c and d, where data from the Convenience kinetics model is used for analysis.

	We did not observe any differences in the degree of identifiability of parameters in $v_5$ between the analyses that used data from either the MWC model or the Convenience kinetics model for $v_3$ (Figure \ref{fig:ident}d and h). The identifiability of parameters based on the first root obtained from the computer algebra system are shown in Figure \ref{fig:ident_root2}c and f.

@@

	\subsection{Rationale for a priori experimental design through practical parameter identification}\label{sec:data_utility}
	%In section \ref{sec:example}, we described how the practical identifiability of parameters in a kinetic model of a metabolic network depended on the enzyme kinetic rate law used to model the fluxes that constitute the network. In this section we look at the rationale for a priori experimental design to collect meaningful data for parameter estimation. 	
	%We also showed how the methodology that we propose can help in determining the nature and type of experiments that one needs to perform to identify and estimate parameters in kinetic models in section \ref{sec:experiment_type}. 	
	Although the need for better parameter estimation methods is well researched, within the kinetic modeling community (\textcolor{red}{include references related to Banga papers on global optimization methods for parameter estimation, Rudiyanto Gunawan, Lennart Ljung and Heijnen papers here}), the need for experimental design methods to satisfy the data needs for parameter estimation has not gained enough traction (\textcolor{red}{include Rudiyanto Gunawan papers, D-optimal design using FIM here}). Hence, we find the need to stress the necessity for experimental design methodologies, in this case, tailored specifically for metabolic network models.

@@

	For $v_1$ (Figure \ref{fig:data_utility}a), data from about 50\% of the 240 combinations of experiments cannot identify any parameter used to model the flux ($V_1^{max}$ and $K_1^{ac}(ne)$). In contrast, data from only about 5\% of the experiments is useless towards identifying any parameter in $v_3$, while there are no useless combinations of experiments from which data cannot be used to identify either $v_2$, or $v_5$. These data are also in agreement with the degree of identifiability observed for parameters in all these fluxes (Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}). Thus, when numerous experiments are possible to obtain steady state data to estimate parameters for a small metabolic network, there is a very high possibility of performing experiments that could provide non-informative data that can result in parameter non-identifiability without a priori experimental design.

>>>>

	\subsection{Identifying parameters in kinetic models of metabolism: an example}\label{sec:example}	
	In this section, we illustrate the proposed methodology step by step to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network} and Section \ref{sec:small-model} of the Supplementary Information). We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) for this demonstration. 
	\begin{align}\tag{S5}
	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
	\end{align}		
	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ for $v_1$, as mentioned in Section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments. So, from the $n_E = 21$ different experiments described in Section \ref{sec:experiments} and Table \ref{tab:pval}, we can choose multiple combinations of $p = 2$ experiments to satisfy the data requirements for identifying $v_1$ i.e., in Equation (\ref{eq:nonlineq}) $j = \{1, 2\}$. We label the available concentrations and fluxes as ${ac}^{(j)}$ and ${v_1}^{(j)}$, respectively. Then, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for $v_1$ as:
	\begin{align*}%\label{eq:nonlin-flux2}
	{v_1}^{(j)} = V_{1}^{max}\frac{ac^{(j)}}{ac^{(j)}+K_{1}^{ac}(ne)} &&  j=\{1, 2\}
	\end{align*}

	Solving this simultaneous system of equations in $\mathbb{R}^2$ using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b). These expressions have the form shown in Equation (\ref{eq:theta-eq}).
	\begin{subequations}\label{eq:v1_par}
		\begin{align}		
		\theta_1 &= V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
		\theta_2 &= K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
		\end{align}
	\end{subequations}
	In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b) for different available experimental data combinations. Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b). 
	We can also apply the proposed methodology to practically identify parameters in $v_1$ under the assumption that the protein concentration for the enzyme \textit{E} is available, in addition to the measured concentration of the metabolites and fluxes. In doing so, we get two expressions similar to the one shown in Equation (\ref{eq:v1_par}) for $k_1^{cat}$ and $K_1^{ac}$. Here, the value of $V_1^{max}$ in Equation (\ref{eq:flux1a}) above is substituted with $V_1^{max} = k_1^{cat}E$ instead. The globally identifiable parameter values for all fluxes modeled as Michaelis-Menten equations ($v_1$ and $v_2$) in provided in Supplementary Figure \ref{fig:parameter_value_proof}a and \ref{fig:parameter_value_proof}b.

	We apply the method described in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, and illustrated above for $v_1$, to other fluxes of the network (Figure \ref{fig:network}) not described using Michaelis-Menten equations ($v_3$ and $v_5$). On applying this method to $v_3$, which is modeled using MWC kinetics, we find that we could not apply our methodology to determine closed form solution for all the parameters in $v_3$. However, when $v_3$ is instead modeled using Convenience kinetics, we find that the parameters are locally identifiable based on the definition we provide in Section \ref{sec:ident}. All three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ used to model $v_3$ have two possible roots. We still could not determine identifiable values for these parameters (Supplementary Figure \ref{fig:parameter_value_v3}) using the constraints on the physical relevance of their values (see Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}). In the box plot in Supplementary Figure \ref{fig:parameter_value_v3}a and b we show all the possible estimated parameter values for both roots of all three parameters modeling $v_3$, when experimental data is generated using the MWC model for $v_3$. Due to the numerous estimated parameter values, we can designate these parameters are practically non-identifiable. Although using the Convenience kinetics model for $v_3$ to generate the in silico experimental data leads to a reduction in the uncertainty of the estimated parameters (Supplementary Figure \ref{fig:parameter_value_v3}c and d), we still end up with non-unique parameter estimates. This leads is to conclude that under the current conditions, flux $v_3$ is structurally locally identifiable and practically non-identifiable.

	However, $K_3^{fdp}$ and $V_3^{max}$ become globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value, and $K_3^{pep}$ and $V_3^{max}$ become globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (Supplementary Figure \ref{fig:parameter_value_v3_var}).

	Through Supplementary Figure \ref{fig:parameter_value_v3_var}, we also want to bring attention to the fact that parameter non-identifiability remains when there is a mismatch between the kinetic rate law used to generate the in silico data, and the kinetic rate law used to identify parameters. For example, in the specific case of $v_3$, we see that globally structurally identifiable parameters $V_3^{max}$ and $K_3^{fdp}$ (when $K_3^{pep}$ is fixed) are practically non-identifiable when data generated from experiments using the MWC kinetics is used to model $v_3$ (Supplementary Figure \ref{fig:parameter_value_v3_var}a). The same is true of globally structurally identifiable parameters $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is fixed (Supplementary Figure \ref{fig:parameter_value_v3_var}b). Instead, when Convenience kinetics is used to generate in silico data, under the aforementioned conditions all the parameters are globally structurally and practically identifiable (Supplementary Figure \ref{fig:parameter_value_v3_var}c and d).

	Similarly, we applied the demonstrated methodology to identify flux $v_5$ that describes a transcription/translation reaction using Hill kinetics, based on only the available metabolomics and the fluxomics information. When we tried identifying the three parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$, we could not obtain closed form solutions for these parameters using the computer algebra system. However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$. Thus, similar to $v_3$ earlier, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in Section \ref{sec:ident}. In this instance, the identification required data from two different experiments. Again, just as we see for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally structurally identifiable. On the other hand, $V_e^{max}$ has only one possible root, thereby it is a globally structurally identifiable parameter. However, only one of the two roots ($K_e^{fdp}(2)$) is practically identifiable based on the physical relevance criteria stated in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b. Thus, enforcing this criteria results in global practical identifiability of both $V_e{max}$ and $K_e^{fdp}$ in $v_5$ (Supplementary Figure \ref{fig:parameter_value_proof}c).

	\textcolor{red}{(should this be moved to discussions?)}So, in summary, we find that the identifiability of parameters in kinetic models of metabolism using steady state information is dependent on the kinetic rate law used to model the fluxes within metabolism. Although in some instances (e.g., $v_5$) local practical identifiability could be resolved to obtain global practical identifiability using constraints on the values of the parameters such that they are physically relevant, this may not be possible in other scenarios (e.g., $v_3$). Moreover, we also see that the structural identifiability of the parameters in any given kinetic rate law model has a bearing on the ability to determine the practical identifiability of parameters using steady state metabolomic, fluxomic and proteomic information. We find that these can sometimes be resolved by reducing the dimension of the parameter space that is being identified: $\theta \in \mathbb{R}^3$ to $\theta \in \mathbb{R}^2$ for both $v_3$ and $v_2$. Additionally, we would also like to point out that discrepancies between in vivo kinetic rate law from which typical experimental data is obtained, and the in vitro rate law used in kinetic models can itself lead to practical parameter non-identifiability or local identifiability. This can lead to uncertainty in parameter estimates made from in vivo experimental data.

@@

	\subsection{Degree of identifiability of different enzyme kinetic rate laws}\label{sec:ident_rate_laws}		
	In Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}, we show the number and percentage of experimental data combinations that are capable of identifying each parameter in each of the four fluxes $v_1$ (Figure \ref{fig:ident}a and e), $v_2$ (Figure \ref{fig:ident}b and f), $v_3$ (Figure \ref{fig:ident}c and g) and $v_5$ (Figure \ref{fig:ident}d and h) belonging to the network shown in Figure \ref{fig:network}. Based on the definition given in Section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter. We see from Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2} that the degree of identifiability of the maximum reaction rates ($V_i^{max}$), whenever they are identifiable for each of the four fluxes, is greater than or equal to the degree of identifiability of the corresponding enzyme binding ($K_i$) constants, or the allosteric activation constant ($K_3^{pep}$), in the respective reaction rate law models. We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience kinetic rate law and $v_5$ is modeled as a Hill equation with inhibition. The fact that in silico experimental data is generated using different kinetic rate law models for $v_3$ (MWC in Figure \ref{fig:ident}a-d vs Convenience kinetics in Figure \ref{fig:ident}e-h) has no bearing on this observation. We accordingly surmise that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.

@@

	The difference in the degrees of identifiability between $v_1$ (Figure \ref{fig:ident}a) and $v_2$ (Figure \ref{fig:ident}b) can be attributed to the ability of data from different combinations of experiments to satisfy the conditions determined for practical identifiability of that parameter. To elaborate, we showed in Equation (\ref{eq:v1_par}) and Section \ref{sec:example} that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them. So, in this instance, recall that the ability of a combination of experiments to be useful is determined by the ability of the input change (changes in the input acetate concentration) to effect a change in the measured value of $v_1$. Similarly, the identification of parameters for $v_2$ (Figure \ref{fig:ident}b) requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}. We elaborate on the type and nature of experiments that are useful and informative for parameter identification in $v_1$ and $v_2$ later in Section \ref{sec:experiment_type}.

	The difference in the degree of identifiability of $v_1$ that comes about from the use of experimental data from different rate laws used to model $v_3$ (Figures \ref{fig:ident}a and e) can be again attributed to the usefulness of experimental data. However, the difference in data usefulness between similar experiments arising due to differences in the kinetic rate law used for $v_3$ could be due to the difference in the dynamics of the network represented by either of the two models. The interaction between different components of the network (Supplementary Figure \ref{fig:network}) as exemplified by the different models plays an important role in determining the dynamic and the steady state response of the network. The difference in dynamics in turn contributes to different experimentally observed values of metabolite concentrations and fluxes (Supplementary Figure \ref{fig:experimental_data_dist}), which in turn changes the usefulness of the given combination of datasets, and its subsequent effect of parameter identifiability.

	Nonetheless, we would like to note that these variations in the dynamics contributed by the difference between the rate laws can be resolved with more information, i.e., experimental proteomic data that provide protein concentrations for different experiments. We earlier saw that relaxing the assumption on the available experiment data, i.e., assuming the availability of protein concentration (\textit{E}), in addition to the metabolite concentrations (\textit{pep} and \textit{fdp}) and fluxes, results in removal of uncertainty and global identifiability of $v_1$ (Supplementary Figure \ref{fig:parameter_value_proof}). Recall that any predicted parameter value for both $k_1^{cat}$ and $K_1^{ac}$ matches with the exact known value irrespective of the model (MWC vs Convenience kinetics) from which experimental data is obtained when enzyme concentrations are included for identifiability (Supplementary Figure \ref{fig:parameter_value_proof}a).  Additionally, we also see an improvement in the degree of identifiability of the corresponding parameters $k_1^{cat}$ and $K_1^{ac}$ (Supplementary Figure \ref{fig:ident_root2}a and b) when enzyme concentrations are available.

	We would also like to point out that a difference in either the ability to estimate true parameter values, or the the identifiability of parameter modeling $v_2$ does not arise due to difference in the kinetic models used to obtain in silico experimental data. This could be because of the fact that any and all available information that is required for the identification of parameters in $v_2$ is encompassed in the steady state data used for identifiability analysis.

	However, unlike $v_2$, we see that the degree of identifiability of both locally identifiable roots of parameters in $v_3$ changes with the change in the data (MWC vs Convenience kinetics) used for the analysis (Figures \ref{fig:ident}c and g and Supplementary Figures \ref{fig:ident_root2}b and e). Despite the difference in the dynamics represented by the different data sets, and its impact on the usefulness of data sets for identifiability purposes that was explained earlier, we could also allude to the structural non-identifiability between the parameters in $v_3$ for the impact that it would have on the practical identifiability of the parameters using in silico data from different sources. In order to ascertain this, we plot the degrees of identifiability of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ under conditions when they are globally identifiable in Supplementary Figure \ref{fig:ident_v3_var}. To recall from Section \ref{sec:example}, $V_3^{max}$ and $K_3^{fdp}$ are globally identifiable when $K_3^{pep}$ is constant, and $V_3^{max}$ and $K_3^{pep}$ are globally identifiable when $K_3^{fdp}$ is known and fixed. Now that structurally non-identifiability between parameters has been alleviated, we see that the uncertainty introduced by the data obtained from the MWC model still results in a degree of identifiability of only about 5\% for $K_3^{fdp}$ (Supplementary Figure \ref{fig:ident_v3_var}a). However, when $K_3^{fdp}$ is held constant all MWC model-based experimental datasets contribute to local practical identifiability of parameters $V_3^{max}$ and $K_3^{pep}$ (Supplementary Figure \ref{fig:ident_v3_var}c), albeit with a bot of uncertainty (Supplementary Figures \ref{fig:ident_v3_var}b). This ties back with our earlier observation that these parameters are not globally practically identifiable from data sets obtained using the MWC model (Supplementary Figure \ref{fig:parameter_value_v3_var}a and b). However, we see from Supplementary Figures \ref{fig:ident_v3_var}c and d that the globally structurally identifiable parameters are also practically identifiable based on the data obtained from the Convenience kinetics model. This can be corroborated by the parameter values shown in Supplementary Figure \ref{fig:parameter_value_v3_var}c and d, where data from the Convenience kinetics model is used for analysis.

	We did not observe any differences in the degree of identifiability of parameters in $v_5$ between the analyses that used data from either the MWC model or the Convenience kinetics model for $v_3$ (Figure \ref{fig:ident}d and h). The identifiability of parameters based on the first root obtained from the computer algebra system are shown in Supplementary Figure \ref{fig:ident_root2}c and f. We use a similar reasoning to that we suggested for identifiability of $v_2$ above to justify our results.

@@

	\subsection{Rationale for a priori experimental design through practical parameter identification}\label{sec:data_utility}
	%In section \ref{sec:example}, we described how the practical identifiability of parameters in a kinetic model of a metabolic network depended on the enzyme kinetic rate law used to model the fluxes that constitute the network. In this section we look at the rationale for a priori experimental design to collect meaningful data for parameter estimation. 	
	%We also showed how the methodology that we propose can help in determining the nature and type of experiments that one needs to perform to identify and estimate parameters in kinetic models in section \ref{sec:experiment_type}. 	
	\textcolor{red}{(should this section moved to SI?)}Although the need for better parameter estimation methods is well researched, within the kinetic modeling community (\textcolor{red}{include references related to Banga papers on global optimization methods for parameter estimation, Rudiyanto Gunawan, Lennart Ljung and Heijnen papers here}), the need for experimental design methods to satisfy the data needs for parameter estimation has not gained enough traction (\textcolor{red}{include Rudiyanto Gunawan papers, D-optimal design using FIM here}). Hence, we find the need to stress the necessity for experimental design methodologies, in this case, tailored specifically for metabolic network models.

@@

	For $v_1$ (Figure \ref{fig:data_utility}a), data from about 50\% of the 240 combinations of experiments cannot identify any parameter used to model the flux ($V_1^{max}$ and $K_1^{ac}(ne)$). In contrast, data from only about 5\% of the experiments is useless towards identifying any parameter in $v_3$, while there are no useless combinations of experiments from which data cannot be used to identify either $v_2$, or $v_5$. These data are also in agreement with the degree of identifiability observed for parameters in all these fluxes (Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}). Thus, when it is possible to obtain steady state data from numerous experiments to estimate parameters for a small metabolic network, there is a very high possibility of performing experiments that could provide non-informative data that can result in parameter non-identifiability without a priori experimental design.
